---
title: "COMPSCI 415.2 Homework 3"
author: "Gregory Spathias"
date: "June 25, 2018"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


Homework Set-up
```{r}
library(tidyverse)
library(ggplot2)
library(mdsr)

```

##**Section 3.8.1: all exercises**
###**3.8.1.1**

The problem with this plot is that it doesn't represent the quantity of points correctly. In order to improve this plot we can add the argument of position = "jitter" to add random noise to each point to more accurately show the quantity of points graphed.

```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) + geom_point(position = "jitter")
```

###**3.8.1.2**
To control the amount of jittering with geom_jitter() you can use the paramters width and height.

###**3.8.1.3**
geom_jitter and geom_count
- Similarities: both of these functions can be used to avoid overplotting in a plot.
- Differences: geom_jitter adds random noise to each point so that points are not overlaid on top of one another. geom_count counts the number of points at each location then maps the count of points to the point area.


###**3.8.1.4**
The default position adjustment is position dodge.

```{r}
ggplot(data = mpg, mapping = aes(x = class, y = hwy)) + geom_boxplot(position = 'dodge')
```


##**Section 3.9.1: #2 and #4 only**
###**3.9.1.2**
labs() modifies the axis, legend, and plot labels. This function allows you to set the titles of a plot, the axis labels, and captions.

###**3.9.1.4**
```{r}
ggplot(data = mpg, mapping = aes(x = cty, y = hwy)) +
  geom_point() + 
  geom_abline() +
  coord_fixed()
```
This plot shows  that city and highway mpg have a strong positive correlation. If a car has a high city mpg then it will have a high highway mpg. The coord_fixed() function forces a specified ratio between the physical representation of data units on the axes. This is important because it represents the number of units on the y-axis equivalent to a change in one unit of x. geom_abline() adds a reference line to a plot where you can specify the slope and intercept. The default is an intercept of 0 and a slop of +1.


##**Section 4.4: #1 and #2 only**
###**4.4.1**
This code doesn't work because in the second my variable the "i" is not actually an i.

###**4.4.2**
Corrected code below
```{r}
library(tidyverse)
library(dplyr)

#changed dota to data
ggplot(data = mpg) + 
  geom_point(mapping = aes(x = displ, y = hwy))

filter(mpg, cyl == 8) #changed fliter to filter, changed "=" to "=="
filter(diamonds, carat > 3) #changed diamond to diamonds
```


##**Section 5.2.4: #1, #3 and #4 only. You will need to install the nycflights13 package and use the flights data.**
```
install.packages("nycflights13")
```
###**5.2.4.1**
**1** 
Flights that had delay of two or more hours.
```
filter(flights, arr_delay >= 120)
```
**2** 
Flights that flew to Houston(IAH or HOU)

```
filter(flights, dest == 'IAH' | dest == 'HOU')
```
**3** 
Flights operated by United, American or Delta

```
filter(flights, carrier %in% c('UA','AA','DL'))
```


**4**
Flights that departed in summer (July, August, and September)

```
filter(flights, month %in% c(7,8,9))
```
**5**
Flights that arrived more than two hours late, but didn’t leave late
```
filter(flights, dep_delay <= 0 & arr_delay > 120)
```
**6**
Flights that were delayed by at least an hour, but made up over 30 minutes in flight
```
filter(flights, dep_delay >= 60 & (dep_delay - arr_delay) > 30)
```

**7**
Flights that departed between midnight and 6am (inclusive)
``` 
filter(flights, dep_time == 2400 | dep_time <= 600)
```


###**5.2.4.3**
```
filter(flights, is.na(dep_time))
```
8,255 flights.
These are most likely flights that were scheduled, but never flew.

###**5.2.4.4**
* NA^0 is not missing because any number to the 0th power is equal to 1.
* NA | True because as long as one condition of an "or" statement is true, the result will be true
* False & NA is not missing because NA is a missing value for the "and" statement ignores it
* NA*0 is missing because you cannot perform a mathematical operation on a missing value.


##**Section 5.4.1: #1 and #3 only**
###**5.4.1.1**

```
select(flights, dep_time, dep_delay, arr_time, arr_delay)
```

```
select(flights, ends_with('delay'))
```

```
select(flights, contains('delay'))
```

```
select(flights, starts_with('dep'), starts_with('arr'))
```


###**5.4.1.3**
The one_of() function selects any variable that matches one of the strings of objects in a list or vector.

In conjunction with 

```
vars <- c("year", "month", "day", "dep_delay", "arr_delay")
```
one_of can be used in a select function to return the columns that equal vars

```
select(flights, one_of(vars))
```
