---
title: "CompSCIX 415.2 Homework 1"
output: html_document
author: Gregory Spathias
date: 6/12/2018
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
<<<<<<< HEAD
My github repository for my assignments can be found at this URL: [https://github.com/GregSpathias/compscix-415-2-assignments](https://github.com/GregSpathias/compscix-415-2-assignments)


```{r load_packages, warning=FALSE, message=FALSE}
library(mdsr)
library(tidyverse)
```

```{r}
data("WorldCities")
```



QUESTION 1: This data set has 23,018 observations and 10 variables. The variable names are code, name, latitude, longitude, country, countryRegion, population, regionCode, region, and date.

```{r}
glimpse(WorldCities)
View(WorldCities)
```

```{r}
WorldCities <-head(WorldCities, 200) # 200 rows
country_col <- WorldCities$country
unique(country_col)
```


QUESTION 2:

```{r}
region_col <- WorldCities$region
unique(region_col)
```

```{r}
country_col <- select(WorldCities, country)
country_col <- WorldCities %>% select(country)
```

first_this %>% then_this %>% then_this

```{r}
WorldCities %>% select(region) %>% head(5)
```

QUESTION 3: 

```{r}
WorldCities %>% select(country) %>% unique()
```

Visualize it

```{r}
WorldCities %>% ggplot(aes(x = population)) + geom_histogram()
```

```{r}
WorldCities%>% ggplot(aes(x = population))+geom_histogram()+xlab('City Population')
```

QUESTION 4:

```{r}
WorldCities%>% ggplot(aes(x = population))+geom_histogram()+xlab('City Population')+ ggtitle('my title')+ylab('Number of Cities')+theme_bw()
```



=======
My Github repository for my assignments can be found at this URL: [https://github.com/GregSpathias/compscix-415-2-assignments](https://github.com/GregSpathias/compscix-415-2-assignments)
>>>>>>> dc8853d0da92da83d940bb16bbef69f6211b01f1
